Problem statement and suggested project outline

### 2011-11-15 background/phd-description1.tex
paper 2
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@34 43ef1190-90fd-4dd7-a769-d0121f02b80e

>>>>

I note that none of the established SPARQL benchmarks, or any other
benchmark I've seen, regard performance (however it is measured) as a
stochastic variable. It clearly is --- other things running on the
benchmark system, randomness in algorithms, unknown data structure,
etc. --- all contribute to random fluctuations in
performance. Consequently, none of the benchmarks use well-established
statistical methodology for hypothesis testing. They should, as
randomness could influence the conclusions drawn. Rectifying this
problem is a first step in the second paper I plan to write.

However, I wish to perform a much deeper investigation in the same
direction. In physical science and engineering, conventional wisdom
has been that you should only vary one variable at a time to study the
effects of that one variable. In medical science, this has been
abandoned several decades ago, thanks to advances in statistics. In
e.g. a case where the researcher administrates different treatments to
terminally ill patients, some of which may be painful or shorten their
lives, experimental economy is extremely important.

In SPARQL benchmarking, the state of the art is that one compiles data
which one hopes is representative for a wide selection of use cases,
and a select few queries. The performance is then measured. Recently,
\cite{Duan:2011:AOC:1989323.1989340} showed that the data used by all
existing benchmarks is far from representative, leaving many open
problems.

With this in mind, I plan to use ideas from statistical experimental
design to investigate whether it is possible to cover the entire
(realistic) parameter space, thus resolving the problem of choosing
hopefully representative data and queries. This would then be the main
topic of paper 2.
